\newglossaryentry{torrent-desc}
{
  name        = {torrent descriptor},
  description = {Metadata about a torrent, consisting of the location of the tracker (its \gls{IPv6} address) and the \gls{torrent-info}.}
}

\newglossaryentry{torrent-info}
{
  name        = {torrent information},
  description = {Metadata describing the file to be distributed by a torrent. This consists of the file's size, the number of pieces that make up the whole file, the size of each piece and the SHA-1 hashes of all pieces. }
}

\newglossaryentry{torrent-info-hash}
{
  name        = {torrent info hash},
  description = {SHA-1 hash of the \gls{torrent-info}. This hash acts as an identifier for the torrent, used in messages by trackers and peers to identify the torrent.}
}

\newglossaryentry{tracker}
{
  name        = {tracker},
  description = {Server tracking the state of the \gls{swarm} of one or more torrents, allowing clients to discover other \glspl{peer} in the swarm.}
}

\newglossaryentry{peer}
{
  name        = {peer},
  description = {A client participating in a torrent's distribution.}
}

\newglossaryentry{swarm}
{
  name        = {swarm},
  description = {The active \glspl{peer} of a torrent.}
}

\newglossaryentry{seed}
{
  name        = {seed},
  description = {\Gls{peer} which has finished downloading the entire distributed file.}
}

\newglossaryentry{seeder}
{
  name        = {seeder},
  description = {See: \gls{seed}}
}

\newglossaryentry{initial-seed}
{
  name        = {initial seed},
  description = {\Gls{seed} that bootstraps the torrent's distribution. This seed has not downloaded the file from other peers, but has been set up by the distributor with the entire file already available.}
}

\newglossaryentry{leecher}
{
  name        = {leecher},
  description = {\Gls{peer} which are still missing parts of the distributed file.}
}
